---
title: "Life in Months"
author: "Your Name"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: "tango"
    theme: "cosmo"
    df-print: paged
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(dplyr) 
library(lubridate)
library(forcats)
library(tidyr)
library(ggplot2)
library(waffle)
```

## Preparating Your Data

First, we are going to set up your data. For `birth_year` put the year you were born and for `birth_month` put the number that corresponds with the month you were born. `last_year` and `last_month` are set to your birth month minus 1 and your birth year plus 80.

```{r}
birth_year <- 1993
birth_month <- 1 #January
last_year <- birth_year + 80
last_month <- birth_month - 1
```

Next, we are going to create the framework for your life in months. All you have to do for the next chunk is run the code - it will generate a data frame that contains all the years and months from your birth until the month and year 80 years after your birth.

```{r}
life_data <- expand_grid(
  month = month.name,
  year = birth_year:last_year) %>%
  mutate(month = fct_relevel(month, month.name)) %>%
  arrange(year, month) %>%
  group_by(year) %>%
  mutate(month_number = row_number()) %>%
  ungroup() %>%
  filter(!(year == birth_year & month_number < birth_month)) %>% 
  filter(!(year == last_year & month_number > last_month))
```

## Defining your Eras

You will adapt the code below to define your eras and colors for each era. Put the "year,month", "era name", "color" - you can use the name of the color or the hex code for the color (use the color picker here to find the hex codes: https://colorpicker.me/).

You do not have to use the eras defined here - you can add, change, or remove lines as needed.

```{r}
eras <- tribble(
~year_month, ~era, ~fill_color,
"1993,1", "Childhood in Tennessee", "#0B4221",  #modify to you
"1998,5", "K - 2nd grade in Oklahoma", "#285B5D",
"2001,6", "3rd - 12th grade in Texas", "#518393",
"2011,8", "BS @ Olivet Nazarene University", "#7DACBC",
"2015,5", "First real jobs @ World Vision & Barna", "#96B0BB",
#you can add more lines and events as you desire
"2016,9", "PhD @ Northwestern University","#C9B7A5" ,
"2021,8", "Became Statistics Professor at APU", "#E9C6A5",
"2024,2", "Unlived Life", "#E4DECE"  #only change color
)
```

Now run the following code - this will add your eras to your life in months.

```{r}
era_names <- paste0(eras$era)
life_data_era <- life_data %>%
  rowwise() %>%
  mutate(year_month = paste0(c(year, month_number), 
                             collapse = ",")) %>%
  ungroup() %>%
  left_join(eras, by = "year_month") %>%
  fill(era, fill_color) %>%
  mutate(fill_color = fct_inorder(fill_color),
         era = fct_inorder(era)) 
```

You can look at the data using the following code:

```{r eval=FALSE}
View(life_data_era)
```

Now, let's create the waffle chart! Run the code to create the graph, but change the title name to match your name:

```{r}
waffle_life<-life_data_era %>%
  count(era) %>% ## the count of each era is the number of months in that era
  ggplot(aes(fill = era, values = n)) +
  waffle::geom_waffle(color = "#F7F7F7", n_rows = 12, size = 1, flip = FALSE, na.rm=FALSE) + ## make each row a year/12 months
  scale_fill_manual(name = "", values = eras$fill_color) +
  coord_equal() +
  theme_enhance_waffle() +
  theme(legend.text = element_text(size = 8),
        legend.position="bottom",
        axis.ticks = element_blank(),
        plot.background = element_rect(fill = "#F7F7F7", 
                                       color = "#F7F7F7")) +
  ggtitle("Dr. Fitz's Life in Months") + #Change the Title Name
  theme(plot.title = element_text(face = "bold", size = 40))
waffle_life
```

If you get an error that "geom_waffle" is not found, you may need to restart you R Session (under Session).

Now we can save the image - Change out your professor's name with your name. You can adjust the ratio for height/width as needed.

```{r}
ggsave("DrFitz_Life_in_Months.png", plot = waffle_life, 
       device = "png", width = 25, height = 15, dpi = 300)
```
