---
title: "Design Task 4: Educational Attainment"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse)
```

## The Data
The Integrated Public Use Microdata Series (IPUMS) collects, preserves and harmonizes census microdata and provides easy access to this data with enhanced documentation. Data from the U.S. Census and a variety of other sources (Current Population Survey; Demographic and Health Surveys; International Surveys) are available at https://www.ipums.org/. In this design task, you will use International Census data from IPUMS-International, but will restrict your attention to the U.S. alone. The data of interest for this task come in two .csv files. A lot of work has already been done to get these .csv files to their current state. 

The `IPUMS-Sex-Edattain.csv`, gives census counts for each country/year/sex/educational attainment combination. The data set includes counts only for those age 25 or older. “Educational attainment” refers to the highest level of education achieved. 


The questions to visualize are as follows:

- How has educational attainment in the U.S. changed over time?   
- Historically, has there been inequity across males and females with regard to the highest level of educational attainment?   
- How has any inequity across sex in educational attainment changed over time, if at all?  

The goal is to make the answers to the three questions as clear as possible in a single visualization. This includes choosing the best visualization, adding labels and annotations, while also choosing colors and styles that are visually pleasing and accessible.

Here is the original data:
```{r}
sexed <- read_csv("./data/IPUMS-Sex-EdAttain.csv")
```

Here are some modifications to the data to help make appropriate comparisons. First, annotate the code to identify what each code is doing.
```{r}
sex_edu_update <- sexed %>% 
  filter(Country == "United States") %>%  #comment
  mutate(EdAttain = factor(EdAttain, 
                           levels = c("Less than primary completed",
                                      "Primary completed",
                                      "Secondary completed", 
                                      "University completed"))) %>% #comment
  group_by(Year, Sex) %>%   #comment
  mutate(TotalYearGender = sum(Count, na.rm = TRUE)) %>%   #comment
  mutate(Proportion_1 = Count/TotalYearGender)  %>%  #comment
  group_by(Year, EdAttain) %>% #comment
  mutate(TotalYearEdu = sum(Count, na.rm = TRUE)) %>% #comment
  mutate(Proportion_2 = Count/TotalYearEdu) #comment
  
```

Why did we perform the calculations above within our data?

> Insert Answer Here




Intepret what `Proportion_1` and `Proportion_2` measure in our data.

> Insert Answer Here



## Visualizations

Now you are going to choose a visualization to modify that you think best helps answer the questions. 

- How has educational attainment in the U.S. changed over time?   
- Historically, has there been inequity across males and females with regard to the highest level of educational attainment?   
- How has any inequity across sex in educational attainment changed over time, if at all?  

Choose the visualization that you think will help you best answer ALL THREE questions at once. Then update the visualization with labels, appropriate colors, annotations, etc. to make the answers to the questions ABSOLUTELY CLEAR. In addition, be sure the data source is identified on the visualization and provide alternative text.


### Questions to Answer

Which visualization did you choose and why?

>Insert Answer Here


How is your choice of visualization better, compared to the other options, at providing answers to all three questions easily within a single visualization?

>Insert Answer Here



Now delete the other visualizations and modify the one you chose with updates, labels, annotations, data sources, etc. to provide clear and concise answers to the three original questions.

Export the image a png file when completed and submit your code, question answers, and png.

Note: If you feel like you might have a better ideas for a visualization other than adapting one of the provided visualizations you are free to create your own instead, but it must meet the same qualifications.


### Visualization 1

```{r}
sex_edu_update %>% 
  ggplot(aes(x = Year, y = Proportion_1, color = Sex)) +
  geom_line() +
  facet_wrap(~EdAttain)

#ggsave()
```



### Visualziation 2

```{r}
sex_edu_update %>% 
  ggplot(aes(x = Year, fill = EdAttain, y = Proportion_1)) +
  geom_density(position = "fill", stat = "identity") +
  facet_wrap(~Sex)

#ggsave()
```


### Visualization 3

```{r}
sex_edu_update %>% 
  ggplot(aes(x = Year, y = Proportion_1, color = EdAttain)) +
  geom_line() +
  facet_wrap(~Sex)

#ggsave()
```

### Visualziation 4

```{r}
sex_edu_update %>% 
  ggplot(aes(x = Year, fill = Sex, y = Proportion_2)) +
  geom_density(position = "fill", stat = "identity") +
  facet_wrap(~EdAttain)

#ggsave()
```