---
title: "Modifying Visualizations of Amounts"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse) #includes ggplot2 + forcats which we will rely on heavily for visualizing amounts
library(patchwork) #for patching together multiple plots
library(cowplot) #for arranging multiple plots
```

## The General Social Survey  
"Since 1972, the General Social Survey (GSS) has provided politicians, policymakers, and scholars with a clear and unbiased perspective on what Americans think and feel about such issues as national spending priorities, crime and punishment, intergroup relations, and confidence in institutions." (http://gss.norc.org/)

We are going to look at a subset of the data, from 2000-2014

```{r}
#| warning: false
#| message: false
gss <- read_csv("./data/gss_subset.csv")
```


## Bar Plot Basics
Let's start with a basic bar plot for Religion

First, let's look at the count/frequency statistics for the religion variable.  

```{r}
gss %>% 
  count(relig)

gss %>% 
  count(relig) %>% 
  mutate(prop = n/sum(n))
```

Here is the default graphic for a barplot of the counts.  Notice, if we use `geom_bar()` it will automatically summarize for us.

```{r}
gss %>% 
  ggplot(aes(x = relig)) +
  geom_bar()
```


Notice that going with just the default is not going to give us a great graph. 

We could switch axis assignments which can help with label readability:

```{r}
gss %>% 
  ggplot(aes(y = relig)) +
  geom_bar()
```

but we need to consider some modifications. 

## Using forcats

The `forcats` package allows us to work directly with categorical variables in ways that make applying Gestalt Principle easier.

Learn more here: https://forcats.tidyverse.org/ 

### Reordering Variable  

We can use `fct_infreq` to apply the principle of simplicity for comparison.  

```{r}
gss %>% 
  ggplot(aes(y = fct_infreq(relig))) +
  geom_bar()
```


We can use `fct_rev` to do the reverse order.
```{r}
gss %>% 
  ggplot(aes(y = fct_rev(fct_infreq(relig)))) +
  geom_bar()
```



### Collapsing Levels  
We can use `fct_collapse` to redefine our factor levels.  Remember - doing this is a form of analysis because you are making some assumptions about what groups go together!  

```{r}

gss %>% 
  mutate(religion = fct_collapse(relig,
                                 Christian = c("Catholic",
                                               "Christian",
                                               "Orthodox-christian",
                                               "Protestant"),
                                 Eastern = c("Buddhism", 
                                             "Hinduism",
                                             "Other eastern"), 
                                 Jewish = c("Jewish"),
                                 Muslim = c("Moslem/islam"),
                                 None = c("None"),
                                 Unknown = c("No answer", 
                                             "Don't know"),
                                 Other = c("Other", 
                                           "Native american",
                                           "Inter-nondenominational"))) %>% 
  ggplot(aes(y =  fct_infreq(religion))) +
  geom_bar() 
```



### Creating Groups 

Use `facet_wrap` to create groups of visualizations based on a specific variable.

```{r}

gss %>% 
  mutate(religion = fct_collapse(relig,
                                 Christian = c("Catholic",
                                               "Christian",
                                               "Orthodox-christian",
                                               "Protestant"),
                                 Eastern = c("Buddhism", 
                                             "Hinduism",
                                             "Other eastern"), 
                                 Jewish = c("Jewish"),
                                 Muslim = c("Moslem/islam"),
                                 None = c("None"),
                                 Unknown = c("No answer", 
                                             "Don't know"),
                                 Other = c("Other", 
                                           "Native american",
                                           "Inter-nondenominational"))) %>% 
  ggplot(aes(y = fct_infreq(religion))) +
  geom_bar() +
  facet_wrap(~marital, nrow = 2)
```


Do we need to do a bit of data cleaning? Should we consider something other than count to display here? Should we switch the variable?

```{r}
gss %>% 
  mutate(religion = fct_collapse(relig,
                                 Christian = c("Catholic",
                                               "Christian",
                                               "Orthodox-christian",
                                               "Protestant"),
                                 Eastern = c("Buddhism", 
                                             "Hinduism",
                                             "Other eastern"), 
                                 Jewish = c("Jewish"),
                                 Muslim = c("Moslem/islam"),
                                 None = c("None"),
                                 Unknown = c("No answer", 
                                             "Don't know"),
                                 Other = c("Other", 
                                           "Native american",
                                           "Inter-nondenominational"))) %>% 
  count(religion, marital) %>% 
  group_by(religion) %>%
  mutate(proportion = n/sum(n)) %>% 
  filter(marital != "No answer") %>% 
  ggplot(aes(y = fct_reorder(marital, proportion),
             x = proportion, fill = marital)) +
  geom_col() +
  facet_wrap(~religion, nrow = 3) +
  guides(fill = "none")


```



### Group 1 visualization

```{r}

```



### Group 2 Visualization

```{r}

```


## More R Practice

For the following 3 tasks, you will try to recreate the graphics on the [slide](https://docs.google.com/presentation/d/1bUAM-vDeBgu0ufvjmkdsxyQC5zzNWSdI/edit?usp=sharing&ouid=106967575594472014629&rtpof=true&sd=true) as precisely as possible.

### Visualization Task 1

Using the `titanic.rds` dataset, recreate the Visualization Task 1 graphic on the slide as precisely as possible.

```{r}
titanic <- read_rds("./data/titanic.rds")
```


*Hints:*

- Create a new variable that uses `died` and `survived` as levels/categories
- Hex colors: `#D55E00D0`, `#0072B2D0` (no alpha is being used)

```{r}

```


### Visualization Task 2

Use the `athletes_dat` dataset --- extracted from `Aus_althetes.rds` --- to recreate the graphic on the slide as precisely as possible. **Create the graphic twice**: once using `patchwork` and once using `cowplot`. Below is some wrangling code to get started.

```{r}
#load data
Aus_athletes <- read_rds("./data/Aus_athletes.rds")
# Get list of sports played by BOTH sexes
both_sports <- Aus_athletes %>%
  # dataset of columns sex and sport 
  # only unique observations
  dplyr::distinct(sex, sport) %>%
  # see if sport is played by one gender or both
  count(sport) %>%
  # only want sports played by BOTH sexes
  filter(n == 2) %>%
  # get list of sports
  pull(sport)

# Process data
athletes_dat <- Aus_athletes %>%
  # only keep sports played by BOTH sexes
  filter(sport %in% both_sports) %>%
  # rename track (400m) and track (sprint) to be track
  # case_when will be very useful with shiny apps
  mutate(
    sport = case_when(
      sport == "track (400m)" ~ "track",
      sport == "track (sprint)" ~ "track",
      TRUE ~ sport
      )
    )
```

*Hints:*

- Build each plot separately
- Bar plot: lower limit 0, upper limit 95
- Bar plot: shift bar labels by 5 units and top justify 
- Bar plot: label size is 5
- Bar plot: `#D55E00D0` & `#0072B2D0` --- no `alpha`
- Scatterplot: `#D55E00D0` & `#0072B2D0` --- no `alpha`
- Scatterplot: filled circle (use shape = 21), with "white" outline; size is 3
- Scatterplot: `rcc` is red blood cell count; `wcc` is white blood cell count
- Boxplot: outline `#D55E00` and `#0072B2`; shading `#D55E0040` and `#0072B240`
- Boxplot: should be made narrower; 0.5
- Boxplot: Legend is inset in top-right corner of bottom plot
- Boxplot: Space out labels `c("female    ", "male")`
- Boxplot: Use `guides(color = guide_legend(override.aes = list(color = NA, fill = c("#D55E00D0", "#0072B2D0"))))` to make legend solid rectangles instead of boxplots

```{r}

```

### Visualization Task 3

Create the following graphic using `patchwork`.

*Hints:*

- Use plots created in exercise 4
- inset theme is classic
  - Useful values: 0, 0.45, 0.75, 1
- plot annotation `"A"`