---
title: "Create a Title for the Project"
output: flexdashboard::flex_dashboard
---

```{r setup, include=FALSE}
library(flexdashboard)
```

Introductory Dashboard
=====================================  

Be sure to include:  
- Data source/descriptions;  
- any prerequisite data cleaning;  
- questions to be asked of the data and answered in subsequent visualizations.  




Page 2 - Add a Title!
=====================================  
You can use different formats to include multiple visualizations or just focus on one visualization.
https://rmarkdown.rstudio.com/flexdashboard/layouts.html#overview
    
Column {data-width=600}
-------------------------------------
    
### Chart 1
    
```{r}
plot(cars)
```
   
Column {data-width=400}
-------------------------------------
   
### Chart 2

```{r}
plot(cars)
```   
 
### Leave the three hashtags and delete text for no title
    
```{r}
plot(cars)
```

Page 3 Title {data-orientation=rows}
=====================================     
   
Row {data-height=600}
-------------------------------------

### Create interesting Headers
And remember you can add text...

```{r}
plot(cars)
```

Row {data-height=400}
-------------------------------------
   
### Chart 2

```{r}
plot(cars)
```   
    
### Chart 3

```{r}
plot(cars)
```