---
title: "Jump Starting a Visualization"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse) #loads ggplot2 and other packages
library(palmerpenguins)
library(esquisse)
```

In the package `palmerpenguins` there is a data set called `penguins`. We will use that.

```{r load-data}
data("penguins")
#View(penguins)
```

```{r first-viz}
ggplot(penguins) +
 aes(x = flipper_length_mm, fill = species) +
 geom_histogram(bins = 30L) +
 scale_fill_viridis_d(option = "plasma", 
                      direction = 1)
```

## Using `esquisse` to get started  

We will be learning the basic structure of `ggplot2` (grammar of graphics plot 2) throughout the semester, but to help you get started we will use two packages to help us out, the first package being `esquisse`. 

Take note of the variable types in the data.  Now let's make a visualization. Be sure to add a title and axis labels as well:

```{r eval = FALSE}
esquisse::esquisser(penguins)
```

Copy the code generated here in the code chunk below and run it:

```{r}

```


## Making the graph your own with `ggThemeAssist`

Highlight the code for your visualization above. Then in the Tools Menu under "Addins" go to "ggplot Theme Assistant". A GUI should appear that allows you to specify text sizes, colors, etc. After you finish the code will be added to the end of your graph!

## Save your graph

After your code for your graph (no + is needed), add a line of code like so:  

`ggsave(filename = "mypenguinplot.png")`

and run the full code chunk. In your project folder you should now see a png file of your graph. Check the box next to the png (if on the server) and then select More >> Export. Upload to Canvas for participation!



