---
title: "Visualizing Economic and Health Indicators"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number-section: false
    highlight: tango
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, comment = "",
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(gghighlight)
library(ggrepel)
library(extrafont) #adds new font options
library(gridExtra) #provides some additional functionality for faceted plots
loadfonts(device = "all") #must use font_import() first (once per device)
theme_set(theme_minimal()) #sets a universal theme for all the plots!
```

## Gapminder Data 

In the gapminder folder, we have five data sets that provide information by country.  For each one, we will have to organize the data from a wide format, with year as the column names, to a long format.

```{r}
child <- read_csv("gapminder/child_mortality_0_5_year_olds_dying_per_1000_born.csv") %>% 
  pivot_longer(cols = -country, names_to = "year", values_to = "child_mortality")

fertility <- read_csv("gapminder/children_per_woman_total_fertility.csv") %>% 
  pivot_longer(cols = -country, names_to = "year", values_to = "child_per")

income <- read_csv("gapminder/income_per_person_gdppercapita_ppp_inflation_adjusted.csv") %>% 
  pivot_longer(cols = -country, names_to = "year", values_to = "gdpPerCap")

lifeExp <- read_csv("gapminder/life_expectancy_years.csv") %>% 
  pivot_longer(cols = -country, names_to = "year", values_to = "lifeExp")

population <- read_csv("gapminder/population_total.csv") %>% 
  pivot_longer(cols = -country, names_to = "year", values_to = "population")
```


We can then join together the data by their common variables, which are year and country:

```{r}
full_country <- population %>% 
  left_join(child, by = c("country", "year")) %>% 
  left_join(lifeExp, by = c("country", "year")) %>% 
  left_join(fertility, by = c("country", "year")) %>% 
  left_join(income, by = c("country", "year")) %>% 
  mutate(year = as.numeric(year)) #year is being treated as a categorical variable, we want it to be numeric
```


You can download additional data at https://www.gapminder.org/data/

## Visualizing History

Use the provided data (or add more information following the template code for combining data files) and create a visualization that looks at some historical event through the specific metrics of interest.  

```{r}

```


